\documentclass{report}
    \title{50003 - Models of Computation - (Dr Raad) Lecture 2}
    \author{Oliver Killane}
    \date{15/10/21}

\input{../50003 common.tex}

\begin{document}
\maketitle
\lectlink{https://imperial.cloud.panopto.eu/Panopto/Pages/Viewer.aspx?id=a3951f13-3a34-44bc-ba10-adc300d0077f}

\section*{Syntax of a while Language}
We can define a simple while language (if, else, while loops) to build programs from \& to analyse.
\begin{center}
	\begin{tabular}{r c l}
		$B \in Bool$ & ::= & $true | false | E = E | E < E | B \& B | \neg B \dots$                   \\
		$E \in Exp$  & ::= & $x | n | E + E | E \times E | \dots$                                     \\
		$C \in Com$  & ::= & $x :=E | if \ B \ then \ C \ else \ C | C;C | skip | while \ B \ do \ C$ \\
	\end{tabular}
\end{center}
Where $x \in Var$ ranges over variable identifiers, and $n \in \mathbb{N}$ ranges over natural numbers.
\\
\\ We can also define simple expressions (\keyword{SimpleExp}) to work on:
\[E \in SimpleExp \ ::= \ n | E + E | E \times E | \dots\]
\subsubsection*{Operational Semantics for SimpleExp}
\begin{itemize}
	\bullpara{Small-Step}{
		Also called structural, gives a method for evaluating an expression step-by-step.
	}
	\bullpara{Big-Step}{
		Also called Natural, ignores intermediate steps and gives result immediately.
	}
\end{itemize}
\subsubsection*{Big Step Semantics of SimpleExp}
The properties OF $\Downarrow$ are:
\begin{itemize}
	\bullpara{Determinacy}{
		For all $E, n_1$ and $n_2$ if $E \Downarrow n_1$ and $E \Downarrow n_2$ then $n_1 = n_2$
	}
	\bullpara{Totality}{
		For all $E$ there exists an $n$ such that $E \Downarrow n$.
		\\
		\\ We can break this with loops in matching, e.g
		\[\bigstepdef{B-NON-TOTAL}{}{true}{true}{}\]
		As a result, on hitting true will not stop.
	}
\end{itemize}
\subsubsection*{Small Step Semantics of SimpleExp}
Given a realtion $\to$ we can define a new relation $\leftarrow *$ such that:
\[E \leftarrow * E' \text{ holds if and only if } E = E' \text{ or there is some finite sequence } E \to E_1 \to E_3 \to \dots \to E_k \to E' \]
\begin{itemize}
	\bullpara{Normal Form}{
		$E$ is in its normal form (irreducable) if there is no $E'$ such that $E \to E'$
		\\
		\\ In \keyword{SimpleExp} the normal form is the natural numbers.
	}
	\bullpara{Determinacy}{
		For all $E, E_1, E_2$ if $E \to E_1$ and $E \to E_2$ then $E_1 = E_2$.
		\\
		\\ There is at most one next step.
	}
	\bullpara{Confluence}{
		For all $E, E_1, E_2$ if $E \to * E_1$ and $E \to * E_2$ then there exists some $E'$ such that $E_1 \to * E'$ and $E_2 \to* E'$.
		\\
		\\ Determinate $\rightarrow$ Confluent.
		\\
		\\ There are several evaluations paths, but they all get the same end result.
	}
	\bullpara{(Strong) Normalisation}{
		There are no infinite sequences of expressions $E_1 \to E_2 \to E_3 \to \dots$ such that for all $i$, $E_i \to E_{i+1}$.
		\\
		\\ Every evaluation path eventually reaches a normal form.
	}
\end{itemize}
Theorem: for all $E, n_1, n_2$, if $E \to * n_1$ and $E \to * n_2$ then $n_1 = n_2$.
\end{document}
